---
title: "R/Basel conference 2023"
---

## Mahalanobis distance

Measures distance between realization of a random variable and a distribution.
If the random variable is restated in the PCA space, then it corresponds to the
euclidian distance in the PCA space. Therefore, it is scale invariant.

```{r}
ma <- cbind(1:6, 1:3)
S <-  var(ma)
mahalanobis(c(0, 0), 1:2, S)
```


```{r}
x <- matrix(rnorm(100*3), ncol = 3)
stopifnot(mahalanobis(x, 0, diag(ncol(x))) == rowSums(x*x))
        ##- Here, D^2 = usual squared Euclidean distances
```


```{r}
Sx <- cov(x)
D2 <- mahalanobis(x, colMeans(x), Sx)
plot(density(D2, bw = 0.5),
     main="Squared Mahalanobis distances, n=100, p=3") ; rug(D2)
qqplot(qchisq(ppoints(100), df = 3), D2,
       main = expression("Q-Q plot of Mahalanobis" * ~D^2 *
                         " vs. quantiles of" * ~ chi[3]^2))
abline(0, 1, col = 'gray')
```

## `dm` package

The dm object interrogates the RDBMS for table and column information and, where implemented, primary and foreign keys. Currently, primary and foreign keys are only available from Postgres and SQL Server.


```{r}
library(dm)
library(RMariaDB)

fin_db <- dbConnect(
  MariaDB(),
  username = 'guest',
  password = 'relational',
  dbname = 'Financial_ijs',
  host = 'relational.fit.cvut.cz'
)

fin_dm <- dm_from_src(fin_db)
fin_dm
```

```{r}
fin_dm %>%
  dm_draw()
```

This is using the `DiagrammeR` package.


## Allowing comments on quarto docs (Hypothes.is)

https://quarto.org/docs/reference/projects/books.html#comments



